Resource Buying Games

### 2017-04-03 paper.tex
added introduction and conclusion

\documentclass{scrartcl}
\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{framed}
\usepackage{todo}
\usepackage{amsmath}
\usepackage[framed]{ntheorem}

@@

@@

\begin{document}
\input{tikz/settings}
\maketitle
\section{Introduction}

@@

The cost that results from a profile $S\in\mathcal{S}$ is covered by all
players. To model this we consider the resource buying game 
\begin{equation*}
  G = \tupel{N, \mathcal{S}\times\mathcal{P}, \pi}
\end{equation*}
where $\mathcal{P} = \times_{i\in N}\mathcal{P}_{i}$ and
$\mathcal{P}_{i} = |\mathbb{R}_{+}^{|E|}|$ to model how much every player
contributes in paying for the cost of every resource. We denote the element
from $\mathcal{P}_{i}$ for $e\in E$ with $p_{i}^{e}$ and call $e\in E$ bought
under $\tupel{S, p}\in \mathcal{S}\times\mathcal{P}$ if
$c_{e}(S)\leq \sum_{i\in N}p_{i}^{e}$. The function $\pi$ gives the payoff for
every player for a strategy profile $\tupel{S, p}$ by
\begin{equation*}
  \pi_{i}(S, p) = \begin{cases}
    \sum_{e\in E}p_{i}^{e} &\text{ if $S_{i}$ is bought under
    $\tupel{S, p}$ (i.e. all $e\in S_{i}$ are bought)}\\
    \infty &\text{ otherwise}
  \end{cases}
\end{equation*}
Thus, every player tries to minimise her personal contribution to the costs
but ensures that her demand is satisfied.
\section{Matroids}
In this section we present the combinatorial structure of Matroids.
\todo{why matroids?}

We use matroids as a generalisation of the concept of linear independence.

@@

Consider the algorithm of Kruskal which computes spanning trees by iteratively
adding edges which do not close a cycle: every added edge works towards a basis
in $\mathcal{S}$ and opens a dimensionality by connecting already computed
components.  When an edge is added the result of the algorithm can only compute
certain spanning trees, namely those which contain all edges added up to this
point. Fixing certain elements in matroids and therefore restricting the set of
\enquote{useable} bases is captured by the \emph{contraction} operation on
matroids.

@@

disregard all bases which use a certain element.

@@

Note that this deletion operation is not necessarily well defined for all
elements of the ground-set, e.g. deletion of an edge in the spanning tree
matroid which is a bridge yields an empty set of bases and thus no valid
matroid. Analogously, any set of edges that, when removed, disconnect the graph
contains at least one element necessarily present in any basis. We call such
sets cuts\footnote{see Figure \ref{fig:cuts} for an illustration} and define
them as follows.

@@

The algorithm we present in the following chooses a basis for every player by
iteratively adding one element from a cut and contracting the matroid to the
added element while deleting all other elements from the cut. Since cuts are
inclusion-wise minimal deleting all but one element from one cut leaves at
least one possible basis for the player available. For every player we maintain
the elements which are chosen for her up to the current state
(\emph{working basis}) and which elements are still available to build her
basis (\emph{ground-set}) and which bases she can still use. Initially the
working basis is empty and the ground-set and available bases are taken from
the matroid of the player. For every resource we maintain the cost the next
added load causes and how much load already is placed upon the
resource(\emph{current load}). Initially for every resource $e$ the current
load is zero and the next load causes $c_{e}(1)$ cost. We can devide the
algorithm into four steps as presented in Figure \ref{alg:main}.
\begin{figure}
  \begin{tabular}{|lll|}
    \hline
    (i) & Init: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item \parbox{0.7\textwidth}{every player: empty working basis,
            complete ground-set, complete set of valid result bases}
          \item every resource: cost for next load, current load
        \end{itemize}
      }\\\hline
    (ii) & Find: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item inclusion-min (necessary cost)-max cut $C$
          \item minimal cost resource $e$ in $C$
          \item owner $i$ of $C$
        \end{itemize}
      }\\\hline
    (iii) & Pick: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item $i$: add $e$ to working basis, contract matroid to $e$
          \item $e$: update marginal cost and increase load
          \item all player: delete elements $C\setminus\set{e}$ from matroid
        \end{itemize}
      }\\\hline
    (iv) & Break: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item exists player with incomplete basis: goto (ii)
          \item otherwise: return chosen bases and payments
        \end{itemize}
      }\\\hline
  \end{tabular}
  \caption{Algorithm to compute a PNE for unweighted matroid games with
    marginally non-increasing cost functions. Note that in step (ii) only
    inclusion-wise minimal cuts are considered (specifically the one for which
    minimal cost of all elements in the cut is maximal over all considered cuts
    is chosen). Restricting the selection to inclusion-wise minimal cuts
    ensures that removing all other elements preserves possible bases for
    \emph{all} players}
  \label{alg:main}
\end{figure}
By induction on the iterations one can prove that the cost of the chosen
element in step (ii) always decreases. This is a key lemma to prove the
correctness of the algorithm. Because, if we assume to have a beneficial
alternative to the computed basis for one player there is an element with
less cost. This alternative element is not part of the selected cut since
it would be removed for all players and thus keep its marginal cost although
only elements with higher costs are removed (by choice of the element in the
cut). Furthermore, the alternative element cannot be available to the player
for which the cut is selected since both elements open the
\enquote{same dimensionality} in contradiction to choosing a cut. Thus, the
alternative element has been removed before which implies less cost in a
previous step to the iteration where the original element was chosen. This
contradicts the monotonicity of costs of considered elements through the
iterations. Thus
\begin{theo}
  The presented algorithm computes a PNE.
\end{theo}

@@

Using this one can derive a series of inequalities to show that every socially
optimal solution satisfies the given characterisation and hence
\begin{theo}
  Every weighted matroid resource buying game with marginally non-increasing
  cost functions posses a PNE.
\end{theo}
For completeness we present - without going into details - an analogous result
for non-decreasing cost functions. Again starting from a socially optimal
configuration one can derive a pricing vector (called marginal cost pricing)
which forms with the given configuration a PNE. Especially this result is
also viable for non-matroid strategy spaces, yielding:
\begin{theo}
  Let $G$ be a weighted resource buying game with non-decreasing marginal costs
  and let $S^{\ast}$ be a socially optimal solution. Then, marginal cost
  pricing induces a PNE.
\end{theo}

@@

two elements from $Y$ to be become available. This can be formally described

@@

in $\tilde{E}\setminus{X_{1}\cup X_{2}\cup Y_{1}\cup Y_{2}}$ to a high constant

@@

>>>>

\documentclass{scrartcl}
\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{framed}
\usepackage{todo}
\usepackage{amsmath}
\usepackage[framed]{ntheorem}
\usepackage[style=numeric]{biblatex}

@@

\addbibresource{bibliography.bib}

@@

\begin{document}

\input{tikz/settings}
\maketitle
\section{Introduction}
In this paper we examine a certain class of games based on congestion models.
Intuitively speaking we consider a set of resources and actors with individual
demands on this set of resources. We model this competing behaviour of actors
in terms of strategic games which are applicable for analysis in regards of
Nash equilibria. Mainly, we are interested in the existence and computability
of pure Nash equilibria (PNE), i.e. a state where each actor has no interest
to uniliterally change her/his behaviour if all other players stick to their
strategy. We examine congestion models with different cost functions of
resource, different structures of resources each player wants to be available
and different demand of players on her/his resources. As main results we prove
the existence of PNE for
\begin{itemize}
  \item matroid games with marginally non-increasing cost functions
  \item congestion games with marginally non-decreasing cost functions
\end{itemize}
In terms of complexity we present an algorithm to compute a PNE for unweighted
(i.e. uniform demands for all players) matroid games with marginally
non-increasing cost functions. Additionally, we construct for non-matroid
structures congestion games with marginally non-increasing cost functions which
specifically do not allow for PNEs. This paper only presents results and
arguments from \cite{main}. Specifically we focus on intuitions and thus
present proofs not in exhaustive detail but try to conway the central ideas.

We start with defining an appropiate congestion model and an associated
strategic game as well as introducing matroids as combinatorial structure
to model desired configurations of resources for players. Following an
algorithm to compute PNEs in unweighted matroid games with marginally
non-increasing cost functions as well as its correctness proof is presented.
Afterwards weighted matroid games with different cost functions are examined
and concludingly, we study non-matroid configuration spaces.

@@

The cost that results from a profile $S\in\mathcal{S}$ is covered by all
players. To model this we consider the resource buying game 
  G = \tupel{N, \mathcal{S}\times\mathcal{P}, \pi}
where $\mathcal{P} = \times_{i\in N}\mathcal{P}_{i}$ and
$\mathcal{P}_{i} = \mathbb{R}_{+}^{|E|}$ to model how much every player
contributes in paying for the cost of every resource. We denote the element
from $\mathcal{P}_{i}$ for $e\in E$ with $p_{i}^{e}$ and call $e\in E$ bought
under $\tupel{S, p}\in \mathcal{S}\times\mathcal{P}$ if
$c_{e}(S)\leq \sum_{i\in N}p_{i}^{e}$. The function $\pi$ gives the payoff for
every player for a strategy profile $\tupel{S, p}$ by
  \pi_{i}(S, p) = \begin{cases}
    \sum_{e\in E}p_{i}^{e} &\text{ if $S_{i}$ is bought under
    $\tupel{S, p}$ (i.e. all $e\in S_{i}$ are bought)}\\
    \infty &\text{ otherwise}
  \end{cases}
Thus, every player tries to minimise her personal contribution to the costs
but ensures that her/his demand is satisfied.
\section{Matroids}
In this section we present the combinatorial structure of matroids
\todo{why matroids?}
which we use as a generalisation of the concept of linear independence.

@@

Consider the algorithm of Kruskal \cite{kruskal} which computes spanning trees
by iteratively adding edges which do not close a cycle: every added edge works
towards a basis in $\mathcal{S}$ and opens a dimensionality by connecting
already computed components. When an edge is added the result of the algorithm
can only compute certain spanning trees, namely those which contain all edges
added up to this point. Fixing certain elements in matroids and therefore
restricting the set of \enquote{useable} bases is captured by the
\emph{contraction} operation on matroids.

@@

disregard all bases which use a certain element (in Kruskal's algorithm upon
adding an edge we can remove all other edges connecting the same components).

@@

In the context of graphs and the spanning tree matroids the contraction and
deletion operation corresponds to the contraction of two nodes connected by
an edge and the deletion of an edge respectively. Note that a deletion
operation is not necessarily well defined for all elements of the ground-set,
e.g. deletion of an edge in the spanning tree matroid which is a bridge yields
an empty set of bases and thus no valid matroid. Analogously, any set of edges
that, when removed, disconnect the graph contains at least one element
necessarily present in any basis. We call such sets cuts\footnote{see Figure
\ref{fig:cuts} for an illustration} and define them as follows.

@@

The algorithm we present in the following chooses a basis for every player by
iteratively adding one element from a cut and contracting the matroid to the
added element while deleting all other elements from the cut. Since cuts are
inclusion-wise minimal deleting all but one element from one cut leaves at
least one possible basis for the player available. For every player we maintain
the elements which are chosen for her up to the current state
(\emph{working basis}) and which elements are still available to build her
basis (\emph{ground-set}) and which bases she can still use. Initially the
working basis is empty and the ground-set and available bases are taken from
the matroid of the player. For every resource we maintain the cost the next
added load causes and how much load already is placed upon the
resource(\emph{current load}). Initially for every resource $e$ the current
load is zero and the next load causes $c_{e}(1)$ cost. We can devide the
algorithm into four steps as presented in Figure \ref{alg:main}.
  \begin{tabular}{|lll|}
    \hline
    (i) & Init: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item \parbox{0.7\textwidth}{every player: empty working basis,
            complete ground-set, complete set of valid result bases}
          \item every resource: cost for next load, current load
        \end{itemize}
      }\\\hline
    (ii) & Find: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item inclusion-min (necessary cost)-max cut $C$
          \item minimal cost resource $e$ in $C$
          \item owner $i$ of $C$
        \end{itemize}
      }\\\hline
    (iii) & Pick: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item $i$: add $e$ to working basis, contract matroid to $e$
          \item $e$: update marginal cost and increase load
          \item all player: delete elements $C\setminus\set{e}$ from matroid
        \end{itemize}
      }\\\hline
    (iv) & Break: &\hspace{-0.6cm}\parbox{0.8\textwidth}{
        \begin{itemize}
          \item exists player with incomplete basis: goto (ii)
          \item otherwise: return chosen bases and payments
        \end{itemize}
      }\\\hline
  \end{tabular}
  \caption{Algorithm to compute a PNE for unweighted matroid games with
    marginally non-increasing cost functions. Note that in step (ii) only
    inclusion-wise minimal cuts are considered (specifically the one for which
    minimal cost of all elements in the cut is maximal over all considered cuts
    is chosen). Restricting the selection to inclusion-wise minimal cuts
    ensures that removing all other elements preserves possible bases for
    \emph{all} players. Also the algorithm terminates in at most $|N|\cdot|E|$
    iterations (each iteration disgards at least one element for one player).
    For a more detailed pseudo-code representation and an description of how to
    compute the minimal cut and the minimal cost element within we refer to
    \cite[Algorithm 1 and 3.3 A Subroutine to Detect $C^{\ast}, e^{\ast}$ and
    $i^{\ast}$]{main}.}
  \label{alg:main}
By induction on the iterations one can prove that the cost of the chosen
element in step (ii) always decreases. This is a key lemma to prove the
correctness of the algorithm. Because, if we assume to have a beneficial
alternative to the computed basis for one player there is an element with
less cost. This alternative element is not part of the selected cut since
it would be removed for all players and thus keep its marginal cost although
only elements with higher costs are removed (by choice of the element in the
cut). Furthermore, the alternative element cannot be available to the player
for which the cut is selected since both elements open the
\enquote{same dimensionality} in contradiction to choosing a cut. Thus, the
alternative element has been removed before which implies less cost in a
previous step to the iteration where the original element was chosen. This
contradicts the monotonicity of costs of considered elements through the
iterations. Thus
  The presented algorithm computes a PNE.

@@

Using this one can derive a series of inequalities to show that every socially
optimal solution satisfies the given characterisation and hence
\begin{theo}
  Every weighted matroid resource buying game with marginally non-increasing
  cost functions posses a PNE.
\end{theo}
For completeness we state - without going into details - an analogous result
for non-decreasing cost functions. Again starting from a socially optimal
configuration one can derive a pricing vector (called marginal cost pricing)
which forms with the given configuration a PNE. Especially this result is
also viable for non-matroid strategy spaces, yielding:
\begin{theo}
  Let $G$ be a weighted resource buying game with non-decreasing marginal costs
  and let $S^{\ast}$ be a socially optimal solution. Then, marginal cost
  pricing induces a PNE.
\end{theo}

@@

two elements from $Y$ to become available. This can be formally described

@@

in $\tilde{E}\setminus(X_{1}\cup X_{2}\cup Y_{1}\cup Y_{2})$ to a high constant

@@

\section{Conclusion}
Among other restrictions we studied unweighted congestion games with marginally
non-increasing cost functions. For these games matroids are precisely the
configuration spaces to guarantee PNEs as well as allow for efficient
computation of one PNE. For weighted matroid games with marginally
non-increasing cost functions the existence of PNEs can be assured. Also for
weighted congestion games with marginally non-decreasing cost functions PNEs
are given. Hereby the structure of strategy spaces is unrestricted.
Concludingly, marginally non-increasing cost functions demand matroid strategy
spaces as combinatorial structure which capture e.g. spanning trees, linear
independence, \dots. On the other hand marginally non-decreasing cost functions
might induce a certain stability in a market by ensuring the existence of PNEs.

\printbibliography
